Energy and exergy analysis of a cruise ship

### 2017-07-19 Manuscript/Manuscript.tex
Changes to manuscript

\author[EPFL]{Francesco Baldi}
\author[Linnaeus]{Fredrik Ahlgren}
\author[DTU]{Tuong-Van Nguyen}
\author[CTU]{Cecilia Gabrielii}
\author[CTU]{Karin Andersson}

\address[EPFL]{Industrial Process Energy Systems Engineering (IPESE), \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, 1950, Sion, Switzerland}
\address[Linnaeus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
\address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
\address[CTU]{Department of Mechanics and Maritime Sciences, Chalmers University of technology, Gothenburg, Sweden}

@@

\begin{align}
	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} = \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \\
	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d} 
\end{align}

@@

The operational data was collected on board from the ships' machine logging and surveillance system. The on board database tool exported all logging points to Excel-97 files, and due to the extensive amount of data points the export was divided in to 15 individual Excel-files consisting of a total 665 MB. The exported raw-data from the ship was over a time span of a full year and in most data-points in 15-min average. The Excel-files were processed in the Pandas library which is a high performance data analysis tool in Python [**INSERT REF TO PANDAS**]. A new structured naming of headers and a consistent time frequency over all data points were created in a HDF5 table database which is the base for the analysis.

All data-points were checked individually by creating a descriptive statistic and a histogram of each data-point. A filter was applied by setting up a pre-defined maximum and minimum value. As we do not know, or had any meaningful way of deriving, each individual sensors placement, sensor type or calibration status all data must be filtered and checked accordingly for outliers or bad data.

@@

The results are presented in three main forms:
\begin{itemize}
	\item Time series (comparing some relevant trips with a daily time scale)
	\item Distribution plot (or box plot...box plots are easier to compare, distribution plots are easier to interpret)
	\item Sankey diagrams
\end{itemize}

The results should be presented both for the full database and for each operational mode (port-sailing-drifting)

Here we can see the typical example (note: the day was not selected as "typical", but simply randomly) of the energy demand during a winter and a summer day (Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}).

@@

@@

Estimated heat demand
Maximum heat demand 
Minimum heat demand

>>>>

\author[EPFL]{Francesco Baldi}
\author[Linneus]{Fredrik Ahlgren}
\author[DTU]{Tuong-Van Nguyen}
\author[CTU]{Cecilia Gabrielii}
\author[CTU]{Karin Andersson}

\address[EPFL]{Industrial Process Energy Systems Engineering (IPESE), \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne, 1950, Sion, Switzerland}
\address[Linneus]{Kalmar Maritime Academy, Linnaeus University, Kalmar, Sweden}
\address[DTU]{Department of Mechanical Engineering, Technical University of Denmark, Lyngby, Denmark}
\address[CTU]{Department of Mechanics and Maritime Sciences, Chalmers University of technology, Gothenburg, Sweden}

@@

\begin{align}
	\sum_{\mathrm{in}} \dot{E}_{\mathrm{in}} &= \sum_{\mathrm{out}} \dot{E}_{\mathrm{out}} +\dot{E}_{d} \\
	\dot{E}_{\mathrm{fuel}} + \dot{E}_{\mathrm{air}} &= \sum_{{\mathrm{waste}}}\dot{E}_{{\mathrm{waste}}}+\dot{E}_{W}+\dot{E}_{Q,\mathrm{heating}}+\dot{E}_{d} 
\end{align}

@@

This section should be written by FA :-)

@@

Here we can see the typical example (note: the day was not selected as "typical", but simply randomly) of the energy demand during a winter and a summer day (Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}). It can be seen clearly how, as expected, the heating demand in winter is much higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduce need for compartment heating. On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/Pie_TotalEnergy}
	\caption{Total energy demand over one year of opeartions}
	\label{fig:Pie_TotalEnergy}
\end{figure}

The average HVAC daily demand, together with the average total electric power demand, is represented in Figure \ref{fig:PelHVAC_vs_time}. As it can be seen, the HVAC demand represents only a minor contribution to the total electric energy demand, as it is only present for a few months during summer, and at rather low power demand. This is not surprising, since it can be observed that (see Figure \ref{fig:Tair_cum}) the air is always below 27 degC, and 90\% of the time below 17 degC.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/PelHVAC_vs_time}
	\caption{HVAC power demand versus total demand. Daily averages for one year of operations}
	\label{fig:PelHVAC_vs_time}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{Figures/Tair_cum}
	\caption{Cumulative distribution of measured ambient air temperatures}
	\label{fig:Tair_cum}
\end{figure}

@@

Looking at Figure \ref{fig:HeatGen_hist} it is possible to appreciate the amount of heat that is recovered from on board available waste heat (cooling water and exhaust gas). It can be observed that the amount of heat recovered from the HT cooling systems is much lower than expected, with an average of around 350 kW. This value is quite unreasonable compared to the values found in the technical specifications from the yard, that go from 700 kW (when only one auxiliary engine is operted) to 4800 kW (when all main engines and two auxiliary engines are operated). 
\\
\begin{figure}
	\centering
	\includegraphics[width=0.99\linewidth]{Figures/HeatGen_hist}
	\caption{Distribution of thermal power generated from on board waste heat recovery}
	\label{fig:HeatGen_hist}
\end{figure}

Looking at the assumptions, this is mostly related to the assumption that the flow of heat recovery water is fixed to 297 m3/h. When the heat demand is medium or low, since the temperature BEFORE the consumers is fixed to 90 degrees, this limits the temperature drop on the consumers, making the maximum heat exchangeable in the HTHR heat exchangers to be very low. This can be seen in Figure \ref{fig:HTHR_T_hist}, where the histograms of the inlet temperature of the HT water and the HR water in the HTHR24 heat exchanger are plotted together. Since the maximum amount of heat that can be exchanged depends directly on this difference, the explanation is clear.

\begin{figure}
	\centering
	\includegraphics[width=0.99\linewidth]{Figures/HTHR_T_hist}
	\caption{Distribution of temperatures at the HTHR (24) inlet}
	\label{fig:HTHR_T_hist}
\end{figure}

The alternative solution would be to assume a constant temperature at the HTHR24 HR water inlet instead. This would, on the one hand, dramatically increase the contribution of the HT water to the heat demand of the ship. On the other hand, there is no information of what this temperature is, and little certainty that this is how the system is regulated.

Figure \ref{fig:Heat_comparison} shows the heat demand calculated based on different values assumed for the HTHR inlet temperature and of the reference outside temperature for the design demand of the HVAC preheater. The values of the average error (calculated both as absolute value and not) are given in Table \ref{tab:Heat_comparison}

\begin{figure}
	\centering
	\includegraphics[width=0.99\linewidth]{Figures/Heat_comparison}
	\caption{Comparison of the calculated steam demand for different values of the fixed HTHR inlet temperature and of the reference temperature for the design HVAC preheater load. The shaded green area represents the time frame when no heating occurs}
	\label{fig:Heat_comparison}
\end{figure}

\begin{table}
	\centering
	\begin{tabular}{cccc}
		\hline 
		\rule[-1ex]{0pt}{2.5ex} $T_{HTHR,in} $ & $T_{OUT,des} $ & $\epsilon$ & $\epsilon_{abs}$ \\ 
		\hline 
		\rule[-1ex]{0pt}{2.5ex} 70 & -20 & 100\% & 80\% \\ 
		\hline 
		\rule[-1ex]{0pt}{2.5ex} 65 & -20 & 76\% & 67\% \\ 
		\hline 
		\rule[-1ex]{0pt}{2.5ex} 60 & -20 & 57\% & 57\% \\ 
		\hline 
		\rule[-1ex]{0pt}{2.5ex} 55 & -20 & 42\% & 49\% \\ 
		\hline 
		\rule[-1ex]{0pt}{2.5ex} 50 & -20 & 28\% & 42\% \\ 
		\hline 
		\rule[-1ex]{0pt}{2.5ex} 70 & -30 & 70\% & 69\% \\ 
		\hline 
		\rule[-1ex]{0pt}{2.5ex} 50 & -30 & 8\% & 35\% \\ 
		\hline 
		\rule[-1ex]{0pt}{2.5ex} $\dot{m}=K$ = k & -30 & 207\% & 147\% \\ 
		\hline 
	\end{tabular} 
	\label{tab:Heat_comparison}
\end{table}
